---
title: "Lab11"
author: "Diego Ibarra"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
header-includes: \usepackage{setspace}\doublespacing
bibliography: References.bib
csl: apa-6th-edition.csl
sansfont: sans
fontsize: 12pt
---

The use of the highlight ('text') will be reserved for denoting code.

To add emphasis to other text, use **bold** or *italics*

You can specify the following

Unordered list item
* Unordered list item

Ordered list item
1. Ordered list item

Website link
[Google](https://www.google.com)

Equation
$A =\pi \times r^{2}$

This is how you waould add H. Wickham as an in-text citation

After a statement:

[@wickham2011ggplot2]

Date only after author name:

[-@wickham2011ggplot2]

Multiple references:

[@wickham2011ggplot2;@wickham2011tourr]

Headings and sub-headings are created by using the pound (#) symbol followed by a space.

# Main Heading
## Sub-heading 1
### Sub-heading 2


```{r, fig.width = 4, fig.height = 3}
#Create dummy data
A <- c("a", "a", "b", "b")
B <- c(5, 10, 15,20)
dataframe <- data.frame(A, B)

#Plot figure
boxplot(B~A, data =  dataframe)
```


```{r}
library(knitr)
kable(dataframe, digits = 2)
```

```{r}
#install.packages(pander)
library(`pander`)
plant <- c("a", "b", "c")
temperature <- c(20, 20, 20)
growth <- c(0.65, 0.95, 0.15)
dataframe <- data.frame(plant, temperature, growth)
emphasize.italics.cols(3)   # Make the 3rd column italics
pander(dataframe)           # Create the table
```


```{r, include = FALSE}
edidiv <- read.csv("edidiv.csv")
```

# Data Exploration

A preliminary investigation into the biodiversity of Edinburgh, using data from the [NBN Gateway](https://data.nbn.org.uk/).






## What is the species richness across taxonomic groups?

A table of species richness:
```{r, results='asis'}
library(dplyr)

richness <- 
  edidiv %>%
  group_by(taxonGroup) %>%
  summarise(Species_richness = n_distinct(taxonName)) 

pander(richness)
```






A barplot of the table above:
```{r, fig.align="center", fig.width=15, fig.height=8}
barplot(richness$Species_richness, 
        names.arg = richness$taxonGroup, 
        xlab="Taxa", ylab="Number of species", 
        ylim=c(0,600)
        ) 
```


### What is the most common species in each taxonomic group?

A table of the most common species:
```{r}
#Create a vector of most abundant species per taxa
max_abund <-
  edidiv %>%
    group_by(taxonGroup) %>%
    summarise(taxonName = names(which.max(table(taxonName))))

#Add the vector to the data frame
richness_abund <-
inner_join(richness, max_abund, by = "taxonGroup")
richness_abund <- rename(richness_abund, Most_abundant =  taxonName, Taxon = taxonGroup)
```

```{r}
richness_abund <- rename(richness_abund, 
                        "Most Abundant" = Most_abundant,
                        "Species Richness" = Species_richness) #Change the column names
emphasize.italics.cols(3) #Make the 3rd column italics
pander(richness_abund) #Create a table
```


